Redshift Evolution of the Type Ia Supernova Population Stretch Distribution

### 2020-02-17 Nicolas.tex
Started YC changes

%\usepackage{hyperref}
\usepackage{txfonts}
\usepackage{amsmath}
\usepackage[colorlinks,breaklinks]{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}

\usepackage{color}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{1, 0.27, 0}{#1}}}

@@

In this paper, we take a step aside to probe the validity of our modeling of the
SN population that we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one young, the former
having on average lower lightcurve stretches and being brighter after
standardization. We indeed use the correlation between the SN age probed with
local specific star formation rate in \cite{rigault2018} and the SN stretch to
model the expected evolution of the underlying SN stretch distribution as a
function of redshift. This modeling relies on three assumptions: (1) there
\nn{are indeed} 2 populations of SNe~Ia; (2) the relative fraction of each of
these populations as a function of redshift follows the age-modeling made in
\citep{rigault2018} and (3) the underlying distribution of stretch for each
age-sample can be modeled and do not significantly drift. This paper aims at
testing this with data from the literature.

@@

The concept of the SNe~Ia age dichotomy arised when the SN~Ia rate has first
been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
have shown that the relative SNe~Ia rate in galaxies could only be explained if
two populations existed, one young, following the host star formation activity,
and one old following the host stellar mass (the so called “prompt and delayed“
or “A+B“ model). In \cite{rigault2018} we used the specific star formation rate
at the SN location (Local sSFR or LsSFR) to classify which are the prompts
(those with a high LsSFR) and which are the delayed (those with low LsSFR).
Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
to be strongly correlated with the SN host properties \citep{hamuy1996,
hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014, kim19}. Following the “A+B“ model and the connection between SN
stretch and host properties, \cite{howell2007} first discussed the potential
redshift drift of the SN stretch distribution. In this paper we revisit this
analysis with the most recent SNe~Ia dataset..

@@

\section{A ''complete'' Sample}
\label{sec:sample}

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SN~Ia that a
magnitude-limited survey will miss are the reddest and the lowest-stretch ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely \nn{\citep[e.g.,][]{kim19}}.

\nn{Assuming} sufficient (and random) spectroscopic follow-up for acquiring
typing and host redshift, the selection effects of magnitude-limited survey
should be negligible bellow a given redshift where even the faintest \nn{SN~Ia}
could be observed. In contrast, targeted surveys have highly complex selection
functions and will be discarded from our analysis. Fortunately, modern SN
cosmology samples such as the pantheon dataset are now dominated by magnitude
limited surveys.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch and color lightcurve parameters
        of SNe~Ia from the SDSS, PS1 and SNLS samples from the pantheon catalog.
        The individual SN data are shown as blue dots and a 2D histogram is
        shown in gray to highlight point density. The ellipse
        ($x_1=\pm3\,\mathrm{;}\,c=\pm0.3$) is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \cite{scolnic2018a}. \nn{The lines represent
        the $(x_1, c)$ evolution for $m = m_{lim}$ in eq.~\ref{eq:m_lim},
    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8\,$mag. The data
    above these lines are those for which $m > m_{lim}$, meaning they can't be
    observed.}}
    \label{fig:maglim}
\end{figure}

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
distributions of SNe~Ia from these surveys, namely: PanStarrs
\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plan with semi-major axis $x_1=3$ and semi-minor axis $c=0.3$ encapsulates
the full distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011}
and \citet{campbell2013} for similar contours, the second using a more
conservative $c=0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014} we can derive the absolute
standardized magnitude at maximum of light along the aforementioned ellipse
given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
\cite{scolnic2018a}. The faintest SN~Ia is that with $x_1=-1.66$ and $c=0.25$
and has an absolute standardized magnitude at peak in Bessel-B band of
$-18.31\,\mathrm{mag}$. Since one need to detect this object typically a week
before and 10 days after peak to build a suitable lightcurve, the limiting
\nn{standardized} absolute magnitude effectively is approximately
$-18.00\,\mathrm{mag}$. Hence, given the $5\sigma$ point source detection
magnitude limit of a magnitude limited survey $m_{lim}$, one can derive the
maximum redshift above which the faintest SNe~Ia will be missed \nn{using the
relation between apparent magnitude, redshift and absolute magnitude given by:}
\begin{equation}\label{eq:m_lim}
    \nn{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\,\text{mag}}
    < m_{lim}}
\end{equation}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a 24.8 mag $5\sigma$
depth\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT
final release website.}}. This converts to a $z_{lim}=0.60$, in perfect
agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}.
Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
agreement with, e.g., Figure~6 of \cite{scolnic2018a}. This figure could also
suggest a more conservative $z_{lim}$ of 0.27; both will be considered as
discussed below.  In similar redshift range, SDSS has a limiting magnitude of
22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to the
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, such that \cite{kessler2009} and
\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$. Following
these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit for
the rest of \nn{this study}. For the rest of the analysis, we will consider both
cases, the \nn{aforementioned} redshift limits as well as more conservative cuts
for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for
PS1 and $z_{lim}=0.55$ for SNLS (following Fig.~14 of \citealt{perrett2010}).

The complete sample selection is summarized table~\ref{tab:sample} and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
We see in this figure that the redshift limits we have selected roughly
correspond to the peak of these histograms\nn{. As the peaks indicate that we
are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
cuts fall near said peaks.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the pantheon dataset. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
full(light) color responds to the conservative(nominal) cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). \nn{Their spectroscopic follow-up were only done on a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
SN population. The SNf data is particularly useful for studying SN property
drift as it enable us to have a large SN~Ia sample at $z<0.1$.  The HST sample
from Pantheon follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{strolger04}, see
table~\ref{tab:sample}. \nn{We present the stretch distribution, redshift histogram
and stretch cdf of these five surveys Fig.~\ref{fig:sample}}.

\begin{table}
    \centering
    \caption{\nn{Source surveys and number of SNe~Ia used in our ''complete''
    sample. Conservative cuts are indicated in brackets.}}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75.pdf}
    \caption{\nn{\textit{Main}: Stretches of each survey in full and empty markers for the
    conservative and nominal cuts, respectively, as a function of redshift. \textit{Right}:
    normed CDF of each survey's stretch distribution. \textit{Top}: combined redshift 
    histogram of all the data, in dark gray for the conservative cuts;
    in light gray for the nominal cuts.}}
    \label{fig:sample}
\end{figure}

@@

%\mri{Kesslet 2009 | "For the SDSS-II and SNLS samples, the cutoff redshifts are
%estimated to be 0.15 and 0.65, respectively."}
% ADDIND: PS cut from pantheon close to 0.27 see Fig. 6 of Scolnic 2018

@@

In~\cite{rigault2018} we presented a modeling of the evolution of the fraction
of prompt and delayed SNe~Ia as a function of redshift following former work on
rates and delay time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
In short, we assumed that the number of prompt SNe~Ia follows the star formation
activity \nn{(SFR)} in the Universe while the number of delayed SNe~Ia follows
the number of Gyr-old stars in the Universe, i.e. the stellar mass \nn{(M)}.
Hence, if we denote $\delta(z)$ (resp. $\psi(z)$) the fraction of young (resp.
old) SNe~Ia in the Universe as a function of redshift, then their ratio
$\delta/\psi$ is expected to follow the evolution of the specific star formation
rate \nn{(SFR/M)} which goes as $(1+z)^{2.8}$ until $z\sim2$
\citep[e.g.,][]{tasca2015}. Hence since \nn{$\delta(0.05) \sim \psi(0.05)$}
\citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
\citep{mannucci2006,rodney2014}, \nn{and $\delta + \psi = 1$,} then we find in
\cite{rigault2018}:
\begin{align}
    \label{eq:delta}
    \delta(z) & = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
    \mathrm{and}\\
    \psi(z) & = \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
\end{align}
where $K=0.87$. This modelisation is comparable to the evolution predicted by
\cite{childress2014} based on SN rates in galaxies depending on their quenching
time as a function of their stellar mass.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
        the local specific star formation rate (LsSFR) for SNf data used in this
        analysis. The color corresponds to the probability for the SNe~Ia to be
        above $\log(\mathrm{LsSFR}=-10.82)$, which defines $p_y$; see
        \cite{rigault2018}. \textit{Right}: histogram of $p_y$-weighted
        marginalization of the SN stretch. On both panel, we illustrate the
        Base model fitted on the SNf data. The young and old population
        contributions are shown in purple and yellow, respectively.}
    \label{fig:stretchlssfr}
\end{figure*}

\nn{In~\cite{rigault2018} we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown in Fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch
as a function of redshift, given our aforementioned model of the evolution of
the fraction of young and old SNe~Ia with redshift, we need to model the SN
stretch distribution for each age subsample. Given the structure of the
stretch--LsSFR scatter plot shown in Fig.~\ref{fig:stretchlssfr}, we define our
model as follows: for the young population the underlying stretch distribution
is modeled as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution
centered on $\mu_1$ with a $\sigma_1$ width; the old population stretch
distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
same as the young population one. \nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift (eq.~\ref{eq:delta}),
and the stretch distribution of both young and old SNe~Ia, our model for the
underlying distribution of stretch $X_1\left(z\right)$ as a function of
redshift is:}
\begin{align}
    \label{eq:stretchz}
    \nn{X_1\left(z \right) =
    } &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
    (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
\end{align}

The estimation of the 5 free parameters
($\theta\equiv{\mu_1,\mu_2,\sigma_1,\sigma_2,a}$) of the model is given by
minimizing a pseudo-$\chi^2$ $\mathcal{L}$ such that:

\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
    \mathrm{d}x^{i}_{1}, \nn{y^i}}\right),
\end{equation}
\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
    $y^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
    respectively. Depending on the data available, there are two ways to fit
these 5 free parameters}.

\subsubsection{\nn{Parameters estimation with LsSFR data}}
\label{sec:modelpy}

\nn{When using SNf data, we choose $y^i$ to be their given $p_y{}^i$, computed
from their LsSFR data (see Fig.~\ref{fig:stretchlssfr}). \nn{It represents the
probability to be above $\log(\mathrm{LsSFR} = -10.82)$.} Thus, we would have
$\mathcal{P}$ from eq.~\ref{eq:likelihood} to be}

>>>>

%\usepackage{hyperref}
\usepackage{txfonts}
\usepackage{amsmath, mathtools}
\usepackage[colorlinks,breaklinks]{hyperref}
\usepackage[normalem]{ulem}
\usepackage{cancel}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=blue}

\usepackage{color}
\newcommand{\mr}[1]{{\textcolor[rgb]{0.60,0.10,0.6}{#1}}}
\newcommand{\mri}[1]{{\textcolor{red}{#1}}}
\newcommand{\nn}[1]{{\textcolor[rgb]{1, 0.27, 0}{#1}}}
\newcommand{\yc}[1]{{\textcolor{blue}{#1}}}

@@

In this paper, we take a step aside to probe the validity of our modeling of the
SN population that we claim to be constituted of two age-populations
\citep{rigault2013,rigault2015,rigault2018}: one old and one young, the former
having on average lower lightcurve stretches and being brighter after
standardization. We indeed use the correlation between the SN age probed with
local specific star formation rate in \cite{rigault2018} and the SN stretch to
model the expected evolution of the underlying SN stretch distribution as a
function of redshift. This modeling relies on three assumptions: (1) there
\nn{are indeed} 2 populations of SNe~Ia; (2) the relative fraction of each of
these populations as a function of redshift follows the age-modeling made in
\citep{rigault2018} and (3) the underlying distribution of stretch for each
age-sample can be modeled and do not significantly drift. This paper aims at
testing this with datasets from the literature.

@@

The concept of the SNe~Ia age dichotomy aroused when the SN~Ia rate has first
been studied. \cite{mannucci2005, scannapieco2005, sullivan2006, aubourg2008}
have shown that the relative SNe~Ia rate in galaxies could only be explained if
two populations existed, one young, following the host star formation activity,
and one old following the host stellar mass (the so called “prompt and delayed“
or “A+B“ model). In \cite{rigault2018} we used the specific Star Formation Rate
at the SN location (Local sSFR or LsSFR) to classify which are the prompts
(those with a high LsSFR) and which are the delayed (those with low LsSFR).
Furthermore, since the first SNe~Ia host analysis, the SN stretch has been known
to be strongly correlated with the SN host properties \citep{hamuy1996,
hamuy2000}, which has been extensively confirmed since \citep[e.g.][]{neill2009,
sullivan2010, lampeitl2010, kelly2010, gupta2011, dandrea2011, childress2013,
rigault2013, pan2014, kim19}. Following the “A+B“ model and the connection between SN
stretch and host properties, \cite{howell2007} first discussed the potential
redshift drift of the SN stretch distribution. In this paper we revisit this
analysis with the most recent SNe~Ia dataset.

@@

\section{Complete Sample Construction}
\label{sec:sample}

We base our analysis on the latest SNe~Ia compilation: the Pantheon catalog from
\cite{scolnic2018a}. A naive approach to test the SN stretch redshift drift
would be to simply compare the observed SN stretch distributions in a few bins
of redshift.  In practice, selection effects are affecting the observed SN
stretch distributions. Indeed, because the observed SN~Ia magnitude correlates
with the lightcurve stretch (and color), the first SN~Ia that a
magnitude-limited survey will miss are the lowest-stretch (and reddest) ones.
Consequently, if selection effects are not accounted for, one might confuse true
population drift with selection effects, and conversely \nn{\citep[e.g.,][]{kim19}}.

\nn{Assuming} sufficient (and random \yc{unbiased?}) spectroscopic follow-up for acquiring
typing and host redshift, the selection effects of magnitude-limited surveys
should be negligible bellow a given redshift where even the faintest \nn{SN~Ia}
can be observed. In contrast, targeted surveys have highly complex selection
functions and will be discarded from our analysis. Fortunately, modern SN
cosmology samples such as the Pantheon dataset are now dominated by magnitude
limited surveys.

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/zmax_maglim_snls.pdf}
    \caption{\textsc{\texttt{SALT2.4}} stretch ($x_1$) and color ($c$) lightcurve parameters
        of SNe~Ia from the SDSS, PS1 and SNLS samples from the Pantheon catalog.
        The individual SNe are shown as blue dots and a 2D histogram is
        shown in gray to highlight point density. The ellipse
        $(x_1=\pm3, c=\pm0.3)$ is displayed, colored by the
        corresponding standardized absolute magnitude using the $\alpha$ and
        $\beta$ coefficients from \cite{scolnic2018a}. \nn{The lines represent
        the $(x_1, c)$ evolution for $m = m_{lim}$ in Eq.~\ref{eq:m_lim},
    for $z$ between $0.60$ and $z=1.00$ using SNLS's $m_{lim}$ of $24.8$~mag. The SNe
    above these lines are those for which $m > m_{lim}$, meaning they cannot be
    observed.}}
    \label{fig:maglim}
\end{figure}

We present in Fig.~\ref{fig:maglim} the lightcurve stretch and color
distributions of SNe~Ia from \nn{the following surveys}: PanStarrs
\citep[PS1][]{rest2014}, the Sloan Digital Sky Survey
\citep[SDSS][]{frieman2008} and the SuperNovae Legacy Survey
\citep[SNLS][]{astier2006}. An ellipse in the \textsc{\texttt{SALT2.4}} $(x_1,
c)$ plan with $x_1 = \pm 3$ and $c = \pm 0.3$ encapsulates
the full distribution \citep{guy2007,betoule2014}; see also \citet{bazin2011}
and \citet{campbell2013} for similar contours, the second using a more
conservative $|c| \leq 0.2$ cut. Assuming the SN absolute magnitude with $x_1=0$ and
$c=0$ is $M_0=-19.36$ \citep{kessler2009,scolnic2014}, we can derive the absolute
standardized magnitude at maximum of light $M = M_0 - \alpha x_1 + \beta c$ along the aforementioned ellipse
given the standardization coefficient $\alpha=0.156$ and $\beta=3.14$ from
\cite{scolnic2018a}: the faintest SN~Ia is that with $(x_1=-1.66, c=0.25)$
and an absolute standardized magnitude at peak in Bessel-B band of
$M^{t_0}_{min} = -18.31$~mag. Since one need to detect this object typically a week
before and 10 days after peak to build a suitable lightcurve, the limiting
\nn{standardized} absolute magnitude effectively is approximately
$M_{min} = -18.00$~mag. Hence, given the \nn{\sout{$5\sigma$ point source detection}}
magnitude limit $m_{lim}$ of a magnitude limited survey, one can derive the
maximum redshift $z_{lim}$ above which the faintest SNe~Ia will be missed \nn{using the
relation between apparent magnitude, redshift and absolute magnitude given by:}
\begin{equation}\label{eq:m_lim}
    \nn{\cancel{m = \mu(z) + \underbrace{M_{min}^{t_0-10}(x_1, c)}_{=-18.00\;\text{mag}}
    < m_{lim}}}
\end{equation}
\begin{equation}\label{eq:m_lim}
    \yc{\mu(z_{lim}) = m_{lim} - M_{min}} \quad\text{min vs. lim???}
\end{equation}

SNLS typically acquires SNe~Ia in the redshift range $0.4<z<0.8$. At these
redshifts the rest-frame Bessel-B band roughly corresponds to the SNLS-$i$
filter that has a $5\sigma$ depth of 
24.8~mag\footnote{\href{https://www.cfht.hawaii.edu/Science/CFHTLS/cfhtlsfinalreleaseexecsummary.html}{CFHT final release website.}}. 
This converts to a $z_{lim}=0.60$, in perfect agreement with \cite{neill2006}, \cite{perrett2010} and \cite{bazin2011}. \nn{Fig.~14 of \citealt{perrett2010}
suggests a lower limit of $z_{lim}=0.55$; both will be considered as discussed below.}

Similarly, PS1 observes SNe~Ia in the range $0.2<z<0.4$, their $g$-band
$5\sigma$ depth is 23.1 mag \citep{rest2014}, which yields to $z_{lim}=0.30$ in
agreement with, e.g., Fig.~6 of \cite{scolnic2018a}. This figure could also
suggest a more conservative $z_{lim}$ of 0.27.

In similar redshift range, SDSS has a limiting magnitude of
22.5 \citep{dilday2008,sako2008}, which would lead to a $z_{lim}=0.24$. However,
the SDSS surveys were more sensitive to limited spectroscopic resources.
\cite{kessler2009} \nn{pointed out} that during year-1 of SDSS, SNe~Ia with
$r-mag<20.5$ were favored for spectroscopic follow up, corresponding to a
redshift cut at $0.15$. For the rest of the SDSS survey, additional
spectroscopic resources were used, and \cite{kessler2009} and
\cite{dilday2008} show a relative completeness up to $z_{lim}=0.2$. Following
these analyses we will use $z_{lim}=0.2$ as the baseline SDSS redshift limit \sout{for
the rest of \nn{this study}}.

For the rest of the analysis, we will consider \nn{\sout{both
cases,}} the \nn{aforementioned} redshift limits as well as more conservative cuts
for each of these surveys, namely $z_{lim}=0.15$ for SDSS, $z_{lim}=0.27$ for
PS1 and $z_{lim}=0.55$ for SNLS.

The complete sample selection is summarized in Table~\ref{tab:sample}, and the
redshift distribution of these three surveys are shown in Fig.~\ref{fig:cuts}.
We see there that the redshift limits we have selected roughly
correspond to the peak of these histograms. \nn{As the peaks indicate that we
are starting to miss the faintest SNe~Ia, it seems logical to have the redshift
cuts fall near said peaks.}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/hist_surveys_cuts_55.pdf}
    \caption{From top to bottom: histogram of SN~Ia redshifts for SDSS, PS1 and
    SNLS from the Pantheon dataset. The colored part of the
distribution shows which SNe~Ia has been kept in our analysis for they are
supposedly free from selection bias (see section~\ref{sec:sample}). The
dark (resp. light) color responds to the conservative (resp. fiducial) selection cut.}
    \label{fig:cuts}
\end{figure}

In addition, we use the SNe~Ia from the Nearby Supernova Factory
\citep[SNfactory][]{aldering2004} published in \cite{rigault2018} and that have
been discovered from non-targeted searches (114 SNe~Ia, see their section~3 and
4.2.2). \nn{Their spectroscopic follow-up were done on a redshift range of
$0.02<z<0.09$~\citep[as in ][]{rigault2018}, while the searches were much
deeper. As such, these SNe~Ia} should also be a random sampling of the underlying
SN population. The SNf sample is particularly useful for studying SN property
drift as it enables us to have a large SN~Ia sample at $z<0.1$.

\nn{Finally, we included t}he HST sample
from Pantheon\nn{, that} follows the same logic of having a search deeper than the
follow-up and we therefore kept it entirely \citep{strolger04}, see
Table~\ref{tab:sample}. \nn{We present the stretch distribution and redshift histogram of these five surveys in Fig.~\ref{fig:sample}}.

\begin{table}
    \centering
    \caption{\nn{Source surveys and number of SNe~Ia used in our ``complete''
    sample. Conservative cuts are indicated in brackets.}}
    \label{tab:sample}
    \begin{tabular}{l c c}
    \hline\hline\\[-0.8em]
        Survey & z$_{lim}$ & N$_{\mathrm{SN}}$ \\[0.15em]
        \hline\\[-0.8em]
        SNf & -- & 114\\[0.30em]
        SDSS & 0.20(0.15) & 167(82)\\[0.30em]
        PS1 & 0.30(0.27) & 160(122)\\[0.30em]        
        SNLS & 0.60(0.55) & 102(78)\\[0.30em]
        HST & -- & 26\\[0.30em]
        \hline\\[-0.8em]
        Total & -- & 569(422)\\[0.30em]
        \hline
    \end{tabular}
\end{table}

\begin{figure}
    \centering
    \includegraphics[width=0.95\linewidth]{Article_figures/stretchs-cut_btw_hist_cumu_75-lb.pdf}
    \caption{\nn{\textit{Main}: SN stretches in each survey as a function of redshift, using solid (resp. empty) markers for the
    conservative (resp. fiducial) selection. \textit{Top}: combined redshift 
    histogram for all surveys, in dark (resp. light) gray for the conservative (resp. fiducial) selection cuts.}}
    \label{fig:sample}
\end{figure}

@@

%\mri{Kesslet 2009 | "For the SDSS-II and SNLS samples, the cutoff redshifts are
%estimated to be 0.15 and 0.65, respectively."}
% ADDIND: PS cut from Pantheon close to 0.27 see Fig. 6 of Scolnic 2018

@@

In~\cite{rigault2018} we presented a modeling of the evolution of the fraction
of prompt and delayed SNe~Ia as a function of redshift following former work on
rates and delay time distributions \citep[e.g.,][]{mannucci2005,
scannapieco2005, sullivan2006, aubourg2008, childress2014, maozmannucci2014}.
In short, we assumed that the number of prompt SNe~Ia follows the star formation
activity \nn{(SFR)} in the Universe while the number of delayed SNe~Ia follows
the number of Gyr-old stars in the Universe, i.e. the stellar mass \nn{(M)}.
Hence, if we denote $\delta(z)$ (resp. $\psi(z) = 1 - \delta(z)$) the fraction of young (resp.
old) SNe~Ia in the Universe as a function of redshift, then their ratio
$\delta/\psi$ is expected to follow the evolution of the specific star formation
rate \nn{(SFR/M)} which goes as $(1+z)^{2.8}$ until $z\sim2$
\citep[e.g.,][]{tasca2015}. Since \nn{$\delta(0.05) \sim \psi(0.05)$}
\citep{rigault2013,rigault2018,wiseman2020}, in agreement with rate expectations
\citep{mannucci2006,rodney2014}, we conclude in
\cite{rigault2018} that
%\begin{align}
%    \label{eq:delta}
%    \delta(z) &= \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}\,
%    \mathrm{and}\\
%    \psi(z) &= \left( K \times (1+z)^{+2.8} +1 \right)^{-1}
%\end{align}
\begin{equation}
    \label{eq:delta}
    \yc{\delta(z) = \left( K^{-1} \times (1+z)^{-2.8} +1 \right)^{-1}}
\end{equation}
with $K=0.87$. This model is comparable to the evolution predicted by
\cite{childress2014} based on SN rates in galaxies depending on their quenching
time as a function of their stellar mass.

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{Article_figures/model_base_hist.pdf}
    \caption{\textit{Main}: \textsc{\texttt{SALT2.4}} lightcurve stretch ($x_1$) as a function of
        the local specific star formation rate (LsSFR) for SNf SNe used in this
        analysis. The color corresponds to the probability $p_y$ for the SNe~Ia to be
        young, i.e. to have $\log\mathrm{LsSFR} \geq -10.82$; see
        \cite{rigault2018}. \textit{Right}: histogram of $p_y$-weighted
        distribution of the SN stretchs. On both panel, we illustrate the
        Base model fitted on the SNf SNe. The young and old population
        contributions are shown in purple and yellow, respectively.}
    \label{fig:stretchlssfr}
\end{figure*}

To model the evolution of the SN stretch
as a function of redshift, given our aforementioned model of the evolution of
the fraction of young and old SNe~Ia with redshift, we need to model the SN
stretch distribution for each age subsample.

\nn{In~\cite{rigault2018}, we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown in Fig.~\ref{fig:stretchlssfr}}. Given the structure of the
stretch-LsSFR scatter plot, our model is defined as follows: 
\begin{itemize}
    \item for the young population the underlying stretch distribution
    is modeled as a normal distribution $\mathcal{N}(\mu_1, \sigma_1^2)$; 
    \item the old population stretch distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
    \mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a Gaussian mixture where one mode is the
    same as the young population one. \yc{explain meaning of $a$}
\end{itemize}
\nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift (Eq.~\ref{eq:delta}),
and the stretch distribution of both young and old SNe~Ia, our model for the
underlying distribution of stretch $X_1(z)$ as a function of
redshift is:}
\begin{align}
    \label{eq:stretchz}
    \nn{X_1(z) = {}}
    \delta(z) &\times \mathcal{N}(\mu_1,\sigma_1^2) + {}\nonumber\\
    (1 - \delta(z)) &\times \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right].
\end{align}

The \yc{maximum-likelihood estimate} of the 5 free parameters
$\theta\equiv({\mu_1,\mu_2,\sigma_1,\sigma_2,a})$ of the model is obtained by
minimizing a pseudo-$\chi^2$ $\mathcal{L}$ defined as:
\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln \prob{x^{i}_{1}}{\theta;
    \mathrm{d}x^{i}_{1}, \nn{y^i}},
\end{equation}
with
\begin{align}
    \label{eq:likelihoodsnf}
    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, y^i} =
    y^i &\times \mathcal{N}\left(x^{i}_{1} \mid \mu_1, \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
        (1-y^i) &\times \nn{\bigg[} 
        a \times \mathcal{N}\left(x^{i}_{1} \mid \mu_1,
        \sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) + {} \nonumber\\
     &(1-a) \times \mathcal{N}\left(x^{i}_{1} \mid \mu_2,
     \sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}\right) \nn{\bigg]}
\end{align}
where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
$y^i$ are the \textsc{\texttt{SALT2.4}} stretch, its associated error 
and the probability that the SN is young, respectively. 
In practice, to ensure fraction $a$ is constrained between 0 and 1,
we fit for $\alpha$ such
that $a=\arctan(\alpha)/\pi+0.5$, which results into asymmetric error on $a$.

Depending on \yc{whether $y^i$ can be 
estimated directly from LsSFR measurements or not}, there are two ways to proceed.

\subsubsection{\yc{With LsSFR measurements}}
\label{sec:modelpy}
